---
title: "Population size experiment"
author: "Katherine G. Skocelas, Austin J. Ferguson, Clifford Bohm, Katherine Perry, Rosemary Adaji, Charles Ofria"
date: "`r Sys.Date()`"
output: bookdown::gitbook
documentclass: book
github-repo: kgskocelas/Primordium
description: "Supplemental material for our ALife 2021 conference submission"
---

# Population size experiment

By default, all populations of contain 200 organisms. 
This experiment tested if increasing the population size to 2,000 organisms has any substantial effect on evolved restraint. 


## Data cleaning

Load necessary libraries

```{r, echo=T, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(ggridges)
library(scales)
library(khroma)
```

Load the data and trim all the unecessay bits (_e.g._, we initially ran sizes 8x8, 1024x1024 but cut them from the paper to make plots easier to read). 
```{r, echo = T}
# Load the data
df = read.csv('../experiments/2021_03_06__pop_size/evolution/data/scraped_evolution_data_200.csv')
df = rbind(df, read.csv('../experiments/2021_03_06__pop_size/evolution/data/scraped_evolution_data_2k.csv'))
# Trim off NAs (artifiacts of how we scraped the data) and trim to only have gen 10,000
df2 = df[!is.na(df$MCSIZE) & df$generation == 10000,]
# Ignore data for size 8x8 and 1024x1024
df2 = df2[df2$MCSIZE != 8 & df2$MCSIZE != 1024,]
```

We group and summarize the data to make to ensure all replicates are present.
```{r}
# Group the data by size and summarize
data_grouped = dplyr::group_by(df2, MCSIZE, POP)
data_summary = dplyr::summarize(data_grouped, mean_ones = mean(ave_ones), n = dplyr::n())
```

We clean the data and create a few helper variables to make plotting easier. 
```{r echo = T}
# Calculate restraint value (x - 60 because genome length is 100 here)
df2$restraint_value = df2$ave_ones - 60
# Make a nice, clean factor for size
df2$size_str = paste0(df2$MCSIZE, 'x', df2$MCSIZE)
df2$size_factor = factor(df2$size_str, levels = c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
df2$size_factor_reversed = factor(df2$size_str, levels = rev(c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024')))
data_summary$size_str = paste0(data_summary$MCSIZE, 'x', data_summary$MCSIZE)
data_summary$size_factor = factor(data_summary$size_str, levels = c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
# Create a map of colors we'll use to plot the different organism sizes
color_vec = as.character(khroma::color('bright')(7))
color_map = c(
  '16x16' =     color_vec[1],
  '32x32' =     color_vec[2],
  '64x64' =     color_vec[3],
  '128x128' =   color_vec[4],
  '256x256' =   color_vec[5],
  '512x512' =   color_vec[6],
  '1024x1024' = color_vec[7]
)
# Set the sizes for text in plots
text_major_size = 18
text_minor_size = 16 
```

## Data integrity check

Now we plot the number of finished replicates for each treatment to make sure all data are present. 
Rows show the number of samples used for fitness. 
Each bar/color shows a different organism size.
```{r echo = F, fig.width=6, fig.height=6}
# Plot the number of replicates for each organism size
ggplot(data_summary, aes(x = size_factor, y = n)) +
  geom_col(aes(fill = size_factor)) +
  geom_text(aes(y = n + 5, label = n)) +
  scale_fill_manual(values = color_map) +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(POP)) +
  xlab('Organism size') +
  ylab('Number of finished replicates')
```


## Plot

Here we plot all the data. 
The figure is split into 6 subplots, each showing a different organism size. 
Inside each subplot, population size is shown on the x-axis.
```{r echo = F, fig.width=6, fig.height=6}
# Plot the evolved restraint buffer for all reps of each org size as boxplots
ggplot(df2[df2$generation == 10000,], aes(x = as.factor(POP), y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Population size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size - 2, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(cols = vars(size_factor)) 
```


## Statistics

The plot shows that increasing population size increases the evolved restraint buffer at organism sizes. 
Further, we see that the same general trend holds at both population sizes, evolved restriant peaks at size 128x128. 

Finally, we treat each organism size as a group and take a Wilcoxon Rank-Sum test between the population sizes.
```{r echo = F, fig.width=6, fig.height=10}
size_vec = c(16, 32, 64, 128, 256, 512)
df_wilcox = data.frame(data = matrix(nrow = 0, ncol = 3))
colnames(df_wilcox) = c('org_size', 'p_value', 'W')
for(size in size_vec){
  res = wilcox.test(
    df2[df2$MCSIZE == size & df2$POP == 200,]$restraint_value, 
    df2[df2$MCSIZE == size & df2$POP == 2000,]$restraint_value, 
    alternative = 'two.sided') 
  df_wilcox[nrow(df_wilcox) + 1,] = c(size, res$p.value, as.numeric(res$statistic)[1])
}
df_wilcox$less_0.01 = df_wilcox$p_value < 0.01
print(df_wilcox)
```

