---
title: "Finite genome -- size 1024x1024 organisms"
author: "Katherine G. Skocelas, Austin J. Ferguson, Clifford Bohm, Katherine Perry, Rosemary Adaji, Charles Ofria"
date: "`r Sys.Date()`"
output: bookdown::gitbook
documentclass: book
github-repo: kgskocelas/Primordium
description: "Supplemental material for our ALife 2021 conference submission"
---

# Baseline: Varying organism size

Here we show all the data for the baseline experiment, where we vary organism size but otherwise all parameters are set to the defaults. 

For this experiment (with all default parameters), we also ran size 8x8 and 1024x1024 organisms.
In the paper, however, we only included sizes from 16x16 to 512x512. 
Size 8x8 organisms are quick to run, but these smaller organisms see the most noise in the fitness data. 
Conversely, size 1024x1024 organisms take so long to run that it was impractical to run them for each experiment. 

Here, we show these results for the baseline experiment, including this additional sizes.
The configuration script and data for the experiment can be found under `2021_02_26__org_sizes/` in the experiments directory of the git repository. 

## Data cleaning
Load necessary libraries
```{r, echo=T, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(ggridges)
library(scales)
library(khroma)
```

Load the data and trim all the unecessay bits (_e.g._, we initially ran sizes 8x8, 1024x1024 but cut them from the paper to make plots easier to read). 
```{r, echo = T}
# Load the data
df = read.csv('../experiments/2021_02_26__org_sizes/evolution/data/scraped_evolution_data.csv')
# Trim off NAs (artifiacts of how we scraped the data) and trim to only have gen 10,000
df2 = df[!is.na(df$MCSIZE) & df$generation == 10000,]
```

We group and summarize the data to make to ensure all replicates are present.
```{r}
# Group the data by size and summarize
data_grouped = dplyr::group_by(df2, MCSIZE)
data_summary = dplyr::summarize(data_grouped, mean_ones = mean(ave_ones), n = dplyr::n())
```

We clean the data and create a few helper variables to make plotting easier. 
```{r echo = T}
# Calculate restraint value (x - 60 because genome length is 100 here)
df2$restraint_value = df2$ave_ones - 60
# Make a nice, clean factor for size
df2$size_str = paste0(df2$MCSIZE, 'x', df2$MCSIZE)
df2$size_factor = factor(df2$size_str, levels = c('8x8', '16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
df2$size_factor_reversed = factor(df2$size_str, levels = rev(c('8x8', '16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024')))
data_summary$size_str = paste0(data_summary$MCSIZE, 'x', data_summary$MCSIZE)
data_summary$size_factor = factor(data_summary$size_str, levels = c('8x8', '16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
# Create a map of colors we'll use to plot the different organism sizes
color_vec = as.character(khroma::color('bright')(7))
color_map = c(
  '8x8' =       '#333333',
  '16x16' =     color_vec[1],
  '32x32' =     color_vec[2],
  '64x64' =     color_vec[3],
  '128x128' =   color_vec[4],
  '256x256' =   color_vec[5],
  '512x512' =   color_vec[6],
  '1024x1024' = color_vec[7]
)
# Set the sizes for text in plots
text_major_size = 18
text_minor_size = 16 
```


## Data integrity check

Now we plot the number of finished replicates for each treatment to make sure all data are present. 
Each bar/color shows a different organism size.
```{r echo = F, fig.width=6, fig.height=6}
# Plot the number of replicates for each organism size
ggplot(data_summary, aes(x = size_factor, y = n)) +
  geom_col(aes(fill = size_factor)) +
  geom_text(aes(y = n + 2, label = n)) +
  scale_fill_manual(values = color_map) +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  xlab('Organism size') +
  ylab('Number of finished replicates')
```


## Aggregate plots

Here we plot all the data at once. 

### Boxplots

```{r echo = F, fig.width=6, fig.height=6}
# Plot the evolved restraint buffer for all reps of each org size as boxplots
ggplot(df2[df2$generation == 10000,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd'))
```

### Raincloud plots

We can plot the same data via raincloud plots.


```{r echo = F, fig.width=6, fig.height=10}
# Plot same data as a raincloud plot
ggplot(df2[df2$generation == 10000,], aes(x = restraint_value, y = size_factor_reversed)) +
  geom_vline(aes(xintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_density_ridges2(aes(fill = size_factor), scale = 0.5, position = position_nudge(y = 0.3)) +
  geom_jitter(aes(color = size_factor), height = 0.18, size = 0.3) + 
  xlab('Evolved restraint buffer') +
  ylab('Organism size') +
  scale_fill_manual(values = color_map) +
  scale_color_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) 
```


## Statistics

First, we perform a Kruskal-Wallis test across all organism sizes to indicate if variance exists. 
If variance exists, we then perform a pairwise Wilcoxon Rank-Sum test to show which pairs of organism sizes significantly differ. 
Finally, we perform Bonferroni-Holm corrections for multiple comparisons. 

```{r}
  res = kruskal.test(df2$restraint_value ~ df2$MCSIZE, df2)
  df_kruskal = data.frame(data = matrix(nrow = 0, ncol = 3))
  colnames(df_kruskal) = c('p_value', 'chi_squared', 'df')
  df_kruskal[nrow(df_kruskal) + 1,] = c(res$p.value, as.numeric(res$statistic)[1], as.numeric(res$parameter)[1])
  df_kruskal$less_0.01 = df_kruskal$p_value < 0.01
  print(df_kruskal)
```


We see that significant variation exists, so we perform pariwise Wilcoxon tests on each to see which pais of sizes are significantly different. 

```{r}
size_vec = c(16, 32, 64, 128, 256, 512)
df_test = df2
df_wilcox = data.frame(data = matrix(nrow = 0, ncol = 5))
colnames(df_wilcox) = c('size_a', 'size_b', 'p_value_corrected', 'p_value_raw', 'W')
for(size_idx_a in 1:(length(size_vec) - 1)){
  size_a = size_vec[size_idx_a]
  for(size_idx_b in (size_idx_a + 1):length(size_vec)){
    size_b = size_vec[size_idx_b]
    res = wilcox.test(df_test[df_test$MCSIZE == size_a,]$restraint_value, df_test[df_test$MCSIZE == size_b,]$restraint_value, alternative = 'two.sided') 
    df_wilcox[nrow(df_wilcox) + 1,] = c(size_a, size_b, 0, res$p.value, as.numeric(res$statistic)[1])
  }
}
df_wilcox$p_value_corrected = p.adjust(df_wilcox$p_value_raw, method = 'holm')
df_wilcox$less_0.01 = df_wilcox$p_value_corrected < 0.01
print(df_wilcox)
```