---
title: "The Evolution of Genetic Robustness for Cellular Cooperation in Early Multicellular Organisms"
author: "Katherine G. Skocelas, Austin J. Ferguson, Clifford Bohm, Katherine Perry, Rosemary Adaji, Charles Ofria"
date: "`r Sys.Date()`"
output: bookdown::gitbook
documentclass: book
github-repo: kgskocelas/Primordium
description: "Supplemental material for the ALife 2022 conference submission"
---

# Germ Mutation Rate Sweep


This was one of the preliminary experiments we conducted to find the default parameters for Primordium.
However, the data shown here were ran after the system was finalized (with new random number seeds). 
There were no qualitative differences from prior results. 

We vary the _germ_ mutation rate, which is the probability that an offspring experiences a mutation to its restraint buffer during organism reproduction. 
The probability of a mutation is per-genome, not per-bit. 
When a germ mutation occurs, only a change of +/-1 is possible in the restraint buffer. 
The final default germ mutation rate was 0.02 (_i.e._, each organism reproduction has a 2% chance of mutation).

The configuration script and data for the experiment can be found under `2021_02_16__germ_mut_fin/` in the experiments directory of the git repository. 


## Data cleaning
Load necessary R libraries
```{r echo=T, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(ggridges)
library(scales)
library(khroma)
```

Load the data and trim include only the final generation data for sizes 16x16 to 512x512.
```{r echo=T, message=F,warning=F}
# Load the data
df = read.csv('../experiments/2021_02_16__germ_mut_fin/evolution/data/scraped_evolution_data.csv')
#df = read.csv('/research/rogue_cell/Primordium/experiments/2021_02_16__germ_mut_fin/evolution/data/scraped_evolution_data.csv')
# Trim off NAs (artifacts of how we scraped the data) and trim to only have gen 10,000
cat(colnames(df), '\n')
df2 = df[!is.na(df$MCSIZE) & df$generation == 10000,]
# Ignore data for size 8x8 and 1024x1024
df2 = df2[df2$MCSIZE != 8 & df2$MCSIZE != 1024,]
```


We group and summarize the data to ensure all replicates are present.
```{r echo = T, message=F,warning=F}
# Group the data by size and summarize
data_grouped = dplyr::group_by(df2, MCSIZE, MUT)
data_summary = dplyr::summarize(data_grouped, mean_ones = mean(ave_ones), n = dplyr::n())
```

Further cleaning of the data plus adding some variables to make plotting easier. 
```{r echo = T, message=F,warning=F}
# Calculate restraint value (x - 60 because genome length is 100 here)
df2$restraint_value = df2$ave_ones - 60
# Make a nice, clean factor for size
df2$size_str = paste0(df2$MCSIZE, 'x', df2$MCSIZE)
df2$size_factor = factor(df2$size_str, levels = c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
df2$size_factor_reversed = factor(df2$size_str, levels = rev(c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024')))
df2$germ_mut_str = paste('GERM MUT', df2$MUT)
df2$mut_factor = factor(df2$MUT, levels = c(0.01, 0.02, 0.05, 0.10, 0.20, 0.50, 1.00))
data_summary$size_str = paste0(data_summary$MCSIZE, 'x', data_summary$MCSIZE)
data_summary$size_factor = factor(data_summary$size_str, levels = c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
data_summary$germ_mut_str = paste('GERM MUT', data_summary$MUT)
data_summary$mut_factor = factor(data_summary$MUT, levels = c(0.01, 0.02, 0.05, 0.10, 0.20, 0.50, 1.00))
# Create a map of colors we'll use to plot the different organism sizes
color_vec = as.character(khroma::color('bright')(7))
color_map = c(
  '16x16' =     color_vec[1],
  '32x32' =     color_vec[2],
  '64x64' =     color_vec[3],
  '128x128' =   color_vec[4],
  '256x256' =   color_vec[5],
  '512x512' =   color_vec[6],
  '1024x1024' = color_vec[7]
)
# Set the sizes for text in plots
text_major_size = 18
text_minor_size = 16 
boxplot_color = '#9ecae1'
```


## Data integrity check

Now we plot the number of finished replicates for each treatment to make sure all data are present. 
Each row shows a different germ mutation rate. 
Each bar/color shows a different organism size.
```{r echo = F, fig.width=6, fig.height=10}
# Plot the number of replicates for each organism size
ggplot(data_summary, aes(x = size_factor, y = n)) +
  geom_col(aes(fill = size_factor)) +
  geom_text(aes(y = n + 4, label = n)) +
  scale_fill_manual(values = color_map) +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) +
  xlab('Organism size') +
  ylab('Number of finished replicates')
```


## Aggregate plots
### Facet by germ  mutation rate

Here we plot all the data at once. 
Each row shows a different germ mutation rate and each boxplot shows a given organism size.

```{r echo = F, fig.width=6, fig.height=10}
# Plot the evolved restraint buffer as boxplots
  # x-axis = organism size
  # y-axis = average evolved restraint buffer for each replicate
  # facet rows = germ mutation rate
ggplot(df2[df2$generation == 10000,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor))
```

Here is the same data, plotted identically other than now each row can have a different y-axis.

```{r echo = F, fig.width=6, fig.height=10}
# Same plot, but with free y-axes between plots
ggplot(df2[df2$generation == 10000,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor), scales = 'free_y')
```


### Facet by organism size

Next, we plot the same data, but this time each row corresponds to a certain organism size while germ mutation rate varies along the x-axis. 

```{r echo = F, fig.width=6, fig.height=10}
# Plot the evolved restraint buffer as boxplots
  # x-axis = germ mutation rate
  # y-axis = average evolved restraint buffer for each replicate
  # facet rows = organism size
ggplot(df2[df2$generation == 10000,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor)) 
```

Again, we plot the same data again, but now the y-axis can change between rows. 

```{r echo = F, fig.width=6, fig.height=10}
# Same plot, but with a free y-axis for each row
ggplot(df2[df2$generation == 10000,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor), scales = 'free_y')
```


## Single organism size plots

Here we plot each organism size independently, with the germ mutation rate on the x-axis. 

### Organism size 16x16

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 16,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor))
```

### Organism size 32x32

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 32,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor))
```

### Organism size 64x64

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 64,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor))
```

### Organism size 128x128

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 128,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor))
```

### Organism size 256x256

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 256,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor))
```

### Organism size 512x512

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 512,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Germ mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor))
```



## Single organism size plots

Similarly, here we plot each germ mutation rate independently, with the organism size on the x-axis. 

### Germ mut. rate 0.01

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 0.01,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```

### Germ mut. rate 0.02

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 0.02,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```

### Germ mut. rate 0.05

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 0.05,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```

### Germ mut. rate 0.1

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 0.1,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```

### Germ mut. rate 0.2

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 0.2,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```

### Germ mut. rate 0.5

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 0.5,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```

### Germ mut. rate 1.0

```{r echo = F, fig.width=6, fig.height=6}
ggplot(df2[df2$generation == 10000 & df2$MUT == 1,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) 
```


## Statistics

Since organism size is our main point of comparison, we calculate stats for each germ mutation rate. 

First, we perform a Kruskal-Wallis test across all organism sizes to indicate if variance exists at that mutation rate. 
If variance exists, we then perform a pairwise Wilcoxon Rank-Sum test to show which pairs of organism sizes significantly differ. 
Finally, we perform Bonferroni-Holm corrections for multiple comparisons. 

```{r}
  mut_vec = c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1)
  df_kruskal = data.frame(data = matrix(nrow = 0, ncol = 4))
  colnames(df_kruskal) = c('germ_mut_rate', 'p_value', 'chi_squared', 'df')
  for(mut_rate in mut_vec){
    df_test = df2[df2$MUT == mut_rate,]
    res = kruskal.test(df_test$restraint_value ~ df_test$MCSIZE, df_test)
    df_kruskal[nrow(df_kruskal) + 1,] = c(mut_rate, res$p.value, as.numeric(res$statistic)[1], as.numeric(res$parameter)[1])
  }
  df_kruskal$less_0.01 = df_kruskal$p_value < 0.01
  print(df_kruskal)
```

We see that significant variation exists within each mutation rate, so we perform pairwise Wilcoxon tests on each to see which pairs of sizes are significantly different. 

```{r}
size_vec = c(16, 32, 64, 128, 256, 512)
mut_vec = c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1)
for(mut_rate in mut_vec){
  df_test = df2[df2$MUT == mut_rate,]
  df_wilcox = data.frame(data = matrix(nrow = 0, ncol = 6))
  colnames(df_wilcox) = c('germ_mut_rate', 'size_a', 'size_b', 'p_value_corrected', 'p_value_raw', 'W')
  for(size_idx_a in 1:(length(size_vec) - 1)){
    size_a = size_vec[size_idx_a]
    for(size_idx_b in (size_idx_a + 1):length(size_vec)){
      size_b = size_vec[size_idx_b]
      res = wilcox.test(df_test[df_test$MCSIZE == size_a,]$restraint_value, df_test[df_test$MCSIZE == size_b,]$restraint_value, alternative = 'two.sided') 
      df_wilcox[nrow(df_wilcox) + 1,] = c(mut_rate, size_a, size_b, 0, res$p.value, as.numeric(res$statistic)[1])
    }
  }
  df_wilcox$p_value_corrected = p.adjust(df_wilcox$p_value_raw, method = 'holm')
  df_wilcox$less_0.01 = df_wilcox$p_value_corrected < 0.01
  print(paste0('Germ mutation rate: ', mut_rate))
  print(df_wilcox)
}
```
