---
title: "Finite genome -- somatic mutation rate sweep"
author: "Katherine G. Skocelas, Austin J. Ferguson, Clifford Bohm, Katherine Perry, Rosemary Adaji, Charles Ofria"
date: "`r Sys.Date()`"
output: bookdown::gitbook
documentclass: book
github-repo: kgskocelas/Primordium
description: "Supplemental material for our ALife 2021 conference submission"
---

# Somatic Mutation Rate Sweep

This experiment was one of the prelimary experiments we conducted to find the default parameters for Primordium. 
Here, we vary the somatic mutation rate, the probability that a cell replication will result in the offspring cell having a different restraint value from its parent. 

We settled on a somatic mutation rate of 0.5 (_i.e._, each cell replication has a 50% chance of mutation).

The configuration script and data for the experiment can be found under `2021_02_27__soma_mut_fin/` in the experiments directory of the git repository. 

## Data cleaning
Load necessary libraries
```{r, echo=T, message=F, warning=F}
library(dplyr)
library(ggplot2)
library(ggridges)
library(scales)
library(khroma)
```

Load the data and trim all the unecessay bits (_e.g._, we initially ran sizes 8x8, 1024x1024 but cut them from the paper to make plots easier to read). 
```{r, echo = T}
# Load the data
df = read.csv('../experiments/2021_02_27__soma_mut_fin/evolution/data/scraped_evolution_data.csv')
# Trim off NAs (artifiacts of how we scraped the data) and trim to only have gen 10,000
df2 = df[!is.na(df$MCSIZE) & df$generation == 10000,]
# Ignore data for size 8x8 and 1024x1024
df2 = df2[df2$MCSIZE != 8 & df2$MCSIZE != 1024,]
```

We group and summarize the data to make to ensure all replicates are present.
```{r}
# Group the data by size and summarize
data_grouped = dplyr::group_by(df2, MCSIZE, CELLMUT)
data_summary = dplyr::summarize(data_grouped, mean_ones = mean(ave_ones), n = dplyr::n())
```

We clean the data and create a few helper variables to make plotting easier. 
```{r echo = T}
# Calculate restraint value (x - 60 because genome length is 100 here)
df2$restraint_value = df2$ave_ones - 60
# Make a nice, clean factor for size
df2$size_str = paste0(df2$MCSIZE, 'x', df2$MCSIZE)
df2$size_factor = factor(df2$size_str, levels = c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
df2$size_factor_reversed = factor(df2$size_str, levels = rev(c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024')))
df2$soma_mut_str = paste('soma CELLMUT', df2$CELLMUT)
df2$mut_factor = factor(df2$CELLMUT, levels = c(0.01, 0.02, 0.05, 0.10, 0.20, 0.50, 1.00))
data_summary$size_str = paste0(data_summary$MCSIZE, 'x', data_summary$MCSIZE)
data_summary$size_factor = factor(data_summary$size_str, levels = c('16x16', '32x32', '64x64', '128x128', '256x256', '512x512', '1024x1024'))
data_summary$soma_mut_str = paste('soma CELLMUT', data_summary$CELLMUT)
data_summary$mut_factor = factor(data_summary$CELLMUT, levels = c(0.01, 0.02, 0.05, 0.10, 0.20, 0.50, 1.00))
# Create a map of colors we'll use to plot the different organism sizes
color_vec = as.character(khroma::color('bright')(7))
color_map = c(
  '16x16' =     color_vec[1],
  '32x32' =     color_vec[2],
  '64x64' =     color_vec[3],
  '128x128' =   color_vec[4],
  '256x256' =   color_vec[5],
  '512x512' =   color_vec[6],
  '1024x1024' = color_vec[7]
)
# Set the sizes for text in plots
text_major_size = 18
text_minor_size = 16 
```

## Data integrity check

Now we plot the number of finished replicates for each treatment to make sure all data are present. 
Each row shows a different somatic mutation rate. 
Each bar/color shows a different organism size.
```{r echo = F, fig.width=6, fig.height=10}
# Plot the number of replicates for each organism size
ggplot(data_summary, aes(x = size_factor, y = n)) +
  geom_col(aes(fill = size_factor)) +
  geom_text(aes(y = n + 4, label = n)) +
  scale_fill_manual(values = color_map) +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor)) +
  xlab('Organism size') +
  ylab('Number of finished replicates')
```

## Aggregate plots
### Facet by somatic mutation rate

Here we plot all the data at once. 
Each row showing a different somatic mutation rate and each boxplot shows a given organism size.
```{r echo = F, fig.width=6, fig.height=10}
# Plot the evolved restraint buffer as boxplots
  # x-axis = organism size
  # y-axis = average evolved restraint buffer for each replicate
  # facet rows = soma mutation rate
ggplot(df2[df2$generation == 10000,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor))
```

Here we plot the same data, only we allow the y-axis to vary between rows. 
```{r echo = F, fig.width=6, fig.height=10}
# Same plot, but with free y-axes between plots
ggplot(df2[df2$generation == 10000,], aes(x = size_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Organism size') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(mut_factor), scales = 'free_y') 
```

### Facet by organism size

Next, we plot the same data, but this time each row corresponds to a certain organism size while somatic mutation rate changes along the x-axis. 
```{r echo = F, fig.width=6, fig.height=10}
# Plot the evolved restraint buffer as boxplots
  # x-axis = soma mutation rate
  # y-axis = average evolved restraint buffer for each replicate
  # facet rows = organism size
ggplot(df2[df2$generation == 10000,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Soma mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor)) 
```

Again, we replot the same data but allow the y-axis to vary between rows.
```{r echo = F, fig.width=6, fig.height=10}
# Same plot, but with a free y-axis for each row
ggplot(df2[df2$generation == 10000,], aes(x = mut_factor, y = restraint_value)) +
  geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(fill = size_factor)) +
  xlab('Soma mutation rate') +
  ylab('Evolved restraint buffer') +
  scale_fill_manual(values = color_map) +
  labs(fill = 'Organism size') +
  theme_light() +
  theme(legend.position = 'none') +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(axis.title = element_text(size = text_major_size)) +
  theme(axis.text = element_text(size = text_minor_size)) +
  theme(legend.title = element_text(size = text_major_size)) +
  theme(legend.text = element_text(size = text_minor_size)) +
  theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
  theme(strip.background = element_rect(fill = '#dddddd')) +
  facet_grid(rows = vars(size_factor), scales = 'free_y') 
```


## Single organism size plots

Here we plot each organism size independently, with the somatic mutation rate on the x-axis. 

### Organism size 16x16 
```{r echo=F, fig.width=6, fig.height=6} 
  ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 16,], aes(x = mut_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Soma mutation rate') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(size_factor)) 
```

### Organism size 32x32
```{r echo=F, fig.width=6, fig.height=6} 
  ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 32,], aes(x = mut_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Soma mutation rate') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(size_factor)) 
```

### Organism size 64x64 
```{r echo=F, fig.width=6, fig.height=6} 
  ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 64,], aes(x = mut_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Soma mutation rate') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(size_factor)) 
```

### Organism size 128x128
```{r echo=F, fig.width=6, fig.height=6} 
  ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 128,], aes(x = mut_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Soma mutation rate') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(size_factor)) 
```

### Organism size 256x256 
```{r echo=F, fig.width=6, fig.height=6} 
  ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 256,], aes(x = mut_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Soma mutation rate') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(size_factor)) 
```

### Organism size 512x512 
```{r echo=F, fig.width=6, fig.height=6} 
  ggplot(df2[df2$generation == 10000 & df2$MCSIZE == 512,], aes(x = mut_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Soma mutation rate') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(size_factor)) 
```


## Single somatic mutation rate plots

Here we plot each somatic mutation rate independently, with organism size varying on the x-axis. 

### Somatic mut. rate 0.01
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 0.01,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```

### Somatic mut. rate 0.02
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 0.02,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```

### Somatic mut. rate 0.05
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 0.05,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```

### Somatic mut. rate 0.1
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 0.1,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```

### Somatic mut. rate 0.2
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 0.2,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```

### Somatic mut. rate 0.5
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 0.5,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```

### Somatic mut. rate 1.0
```{r echo=F, fig.width=6, fig.height=6}
  ggplot(df2[df2$generation == 10000 & df2$CELLMUT == 1,], aes(x = size_factor, y = restraint_value)) +
    geom_hline(aes(yintercept = 0), alpha = 0.5, linetype = 'dashed') +
    geom_boxplot(aes(fill = size_factor)) +
    xlab('Organism size') +
    ylab('Evolved restraint buffer') +
    scale_fill_manual(values = color_map) +
    labs(fill = 'Organism size') +
    theme_light() +
    theme(legend.position = 'none') +
    theme(panel.grid.major.x = element_blank()) +
    theme(panel.grid.minor.x = element_blank()) +
    theme(axis.title = element_text(size = text_major_size)) +
    theme(axis.text = element_text(size = text_minor_size)) +
    theme(axis.text.x = element_text(angle = 60, vjust = 0.8, hjust = 0.8)) +
    theme(legend.title = element_text(size = text_major_size)) +
    theme(legend.text = element_text(size = text_minor_size)) +
    theme(strip.text = element_text(size = text_minor_size, color = '#000000')) +
    theme(strip.background = element_rect(fill = '#dddddd')) +
    facet_grid(rows = vars(mut_factor))
```


## Statistics

Since organism size is our main point of comparison, we calculate stats for each somatic mutation rate. 

First, we perform a Kruskal-Wallis test across all organism sizes to indicate if variance exists at that mutation rate. 
If variance exists, we then perfrm a pairwise Wilcoxon Rank-Sum test to show which pairs of organism sizes significantly differ. 
Finally, we perform Bonferroni-Holm corrections for multiple comparisons. 

```{r}
  mut_vec = c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1)
  df_kruskal = data.frame(data = matrix(nrow = 0, ncol = 4))
  colnames(df_kruskal) = c('soma_mut_rate', 'p_value', 'chi_squared', 'df')
  for(mut_rate in mut_vec){
    df_test = df2[df2$CELLMUT == mut_rate,]
    res = kruskal.test(df_test$restraint_value ~ df_test$MCSIZE, df_test)
    df_kruskal[nrow(df_kruskal) + 1,] = c(mut_rate, res$p.value, as.numeric(res$statistic)[1], as.numeric(res$parameter)[1])
  }
  df_kruskal$less_0.01 = df_kruskal$p_value < 0.01
  print(df_kruskal)
```

We see that significant variation exists within each mutation rate, so we perform pariwise Wilcoxon tests on each to see which pais of sizes are significantly different. 

```{r}
size_vec = c(16, 32, 64, 128, 256, 512)
mut_vec = c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1)
for(mut_rate in mut_vec){
  df_test = df2[df2$CELLMUT == mut_rate,]
  df_wilcox = data.frame(data = matrix(nrow = 0, ncol = 6))
  colnames(df_wilcox) = c('mut_rate', 'size_a', 'size_b', 'p_value_corrected', 'p_value_raw', 'W')
  for(size_idx_a in 1:(length(size_vec) - 1)){
    size_a = size_vec[size_idx_a]
    for(size_idx_b in (size_idx_a + 1):length(size_vec)){
      size_b = size_vec[size_idx_b]
      res = wilcox.test(df_test[df_test$MCSIZE == size_a,]$restraint_value, df_test[df_test$MCSIZE == size_b,]$restraint_value, alternative = 'two.sided') 
      df_wilcox[nrow(df_wilcox) + 1,] = c(mut_rate, size_a, size_b, 0, res$p.value, as.numeric(res$statistic)[1])
    }
  }
  df_wilcox$p_value_corrected = p.adjust(df_wilcox$p_value_raw, method = 'holm')
  df_wilcox$less_0.01 = df_wilcox$p_value_corrected < 0.01
  print(paste0('Somatic mutation rate: ', mut_rate))
  print(df_wilcox)
}
```

